Visualising metagenome differences

### 2023-11-21 analysis/metagenome.Rmd
Final figures and changes

```{r fig.height=6}
plot_orders_gtdb <- function(amph_pos_tax, group_remainder = TRUE) {
  temp <- amph_pos_tax |> 
    group_by(sample, Tissue, Sample_species) |> 
    mutate(Sample_species = trunc_str(Sample_species)) |> 
    mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
    mutate(order = str_replace_all(order, 'o__','')) |> 
    count(order) 
  if(group_remainder) {
    temp <- temp |> 
    mutate(counted_orders = case_when(n == 1 ~ 'Remainder',
                                      TRUE ~ order))
  } else {
      temp <- temp |> 
        mutate(counted_orders = order)
  }
  temp <- temp |> 
    select(-order) |> 
    group_by(sample, Tissue, Sample_species, n, counted_orders) |> 
    summarise(n = sum(n)) |> 
    filter(counted_orders != '' )

@@

>>>>

```{r fig.height=6}
plot_orders_gtdb <- function(amph_pos_tax, group_remainder = TRUE) {
  temp <- amph_pos_tax |> 
    group_by(sample, Tissue, Sample_species) |> 
    mutate(Sample_species = trunc_str(Sample_species)) |> 
    mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
    mutate(order = str_replace_all(order, 'o__','')) |> 
    count(order, .drop = FALSE) 
  if(group_remainder) {
    temp <- temp |> 
    mutate(counted_orders = case_when(n == 1 ~ 'Remainder',
                                      TRUE ~ order))
  } else {
      temp <- temp |> 
        mutate(counted_orders = order)
  }
  temp <- temp |> 
    select(-order) |> 
    group_by(sample, Tissue, Sample_species, n, counted_orders) |> 
    summarise(n = sum(n)) |> 
    filter(counted_orders != '' )

@@

Let's chekc for nitrogen fixation genes (*nif*) - hypothesis is that Amphibolis roots contain more of these

```{r}
genessplit |> filter(str_detect(gene, 'nifH') | str_detect(gene, 'acdS') | str_detect(gene, 'sctC')) |> 
  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ '*nifH*',
                                str_detect(gene, 'acdS') ~ '*acdS*',
                                str_detect(gene, 'sctC') ~ '*sctC*')) |>
  #group_by(Sample_species, Tissue, Sample_replicate, MAG) |> 
  #summarise(gene_class = paste0(sort(unique(gene_class)), collapse = ' & ')) |>
  group_by(Sample_species, Tissue, Sample_replicate) |> 
  count(gene_class) |> 
  mutate(Sample_species = trunc_str(Sample_species)) |> 
  mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
  ggplot(aes(x=Tissue, y=n, fill=gene_class)) + 
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(~Sample_species) + 
  geom_point(position = position_jitterdodge(), aes(color=gene_class)) +
  coord_flip() +
  theme_minimal() +
  labs(color='MAG gene content', fill='MAG gene content') +
  theme(strip.text.x =  element_markdown(),legend.text = element_markdown()) +
  ylab('Count of MAGs')

Can we make a simple association test?

```{r}
tmp <- genessplit |> filter(str_detect(gene, 'nifH') | str_detect(gene, 'acdS') | str_detect(gene, 'sctC')) |> 
  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ '*nifH*',
                                str_detect(gene, 'acdS') ~ '*acdS*',
                                str_detect(gene, 'sctC') ~ '*sctC*')) |> 
  group_by(Sample_species, Sample_replicate, Tissue) |> 
  count(gene_class) |> 
  mutate(Sample_species = factor(Sample_species, levels = c('Posidonia australis', 'Amphibolis antarctica'))) 
summary(lm(n ~ gene_class + Sample_species + Sample_replicate + Tissue, data=tmp))
```

OK! So we have an association with gene_class, which is obvious, but we also have an association with species (great!) and root-tissue (also great!)
